---
title: "HEAL_CSCS_2025_patterns_of_social_connection"
author: "Jorge Andres Delgado-Ron"
date: "`r Sys.Date()`"
output: html_document
---

Step 0. Loading packages, opening dataset, and metadata

```{r setup}
# This pakcage runs the script below for all future packages using p_load()
if (!("pacman" %in% installed.packages()[, "Package"])) {
	# Install the 'pacman' package if it's not installed
	install.packages("pacman")
}

library("pacman")
p_load(here)
p_load(dplyr)
p_load(DataExplorer)
p_load(tableone)

load(here("cscs","public_data","CSCS2025_full_cleaned_deidentified_data_and_metadata.RData"))
set.seed(986)

```

Custom functions
```{r}

# Input variable as text
# Returns the unique label(s) and which data frame(s) contain it.

which_dataset <- function(variable){
levels(data$DATASET) <- c("baseline", "baseline", "cohort22", "cross22", "cohort23", "cross23", "cohort24", "cross24")
x <- data %>% select(DATASET, !!sym(variable)) %>% group_by(DATASET) %>% summarise(result = !all(is.na(!!sym(variable)))) %>% select(DATASET, result)  %>% arrange(as.character(DATASET))
y <- unique(labels$text[labels$new_var==variable])
z <- list(y,x)
return(z)
}

# Input list of variables
# Returns a dataframe with variables as rows and data frames as columns

which_datasets_group <- function(list_of_vars){
df_temp <- which_dataset(list_of_vars[1])[[2]]
colnames(df_temp)[2] <- list_of_vars[1]

for(i in 2:length(list_of_vars)){
x <-  which_dataset(list_of_vars[i])[[2]]
colnames(x)[2] <- list_of_vars[i]
df_temp <- merge(df_temp,x, by="DATASET")
}
rownames(df_temp) <- df_temp$DATASET

df <- as.data.frame(t(df_temp)) %>% mutate(variable_name=rownames(.)) %>% relocate(variable_name, .before = baseline) %>% slice(-1) %>% arrange(as.character(variable_name))
rownames(df) <-NULL
return(df)
}


```

Project-specific data cleaning
```{r}
data$CONNECTION_social_num_coworkers_and_classmates_p7d_grouped[is.na(data$CONNECTION_social_num_coworkers_and_classmates_p7d_grouped) & data$DEMO_student=="No" & data$WORK_hours_per_week==0] <- "None (0 People)"
data$CONNECTION_social_time_coworkers_and_classmates_p7d_grouped[is.na(data$CONNECTION_social_time_coworkers_and_classmates_p7d_grouped) & data$DEMO_student=="No" & data$WORK_hours_per_week==0] <- "No time"
data$CONNECTION_social_days_coworkers_and_classmates_p7d_grouped[is.na(data$CONNECTION_social_days_coworkers_and_classmates_p7d_grouped) & data$DEMO_student=="No" & data$WORK_hours_per_week==0] <- "None (0 Days)"

data$WORK_hours_per_week[data$DEMO_unable_to_work=="Unable to work" & is.na(data$WORK_hours_per_week) & !is.na(data$DEMO_unable_to_work)] <- 0
data$WORK_hours_per_week[data$DEMO_retired=="Retired" & is.na(data$WORK_hours_per_week) & !is.na(data$DEMO_retired)] <- 0
```

Step 2. Identifying variables 

In the dataset we have a series of variables capturing patterns of social connection:

•	CONNECTION_activities_...
•	CONNECTION_social_num_...
•	CONNECTION_social_time_...
•	CONNECTION_social_days_...


These variables are part of the **core survey**, and some appear in both the **cross-section** and **cohort** datasets.

### CONNECTION_activities Suffixes

There are three suffixes used for the `CONNECTION_activities` variables:

- **`_last`**  
  - **Question:** *"When was the last time you..."*  
  - **Applies to:** 2022 cross-section only

- **`_p3m`**  
  - **Question:** *"In the PAST THREE MONTHS, how often have you..."*  
  - **Applies to:** All other datasets

- **`_pm`**  
  - **Derived variable**  
  - **Purpose:** Attempts to capture both `_last` and `_p3m` within the **past month**  
  - **Values:** `"Yes"` | `"No"`

### Response Interpretation

- **"Yes" includes:**
  - For `_last`:  
    *"In the past... month, two weeks, week, 2 or 3 days, yesterday, earlier today"*
  - For `_p3m`:  
    *"A few times a month, monthly, weekly, a few times a week, daily"*

- **"No" includes:**
  - For `_last`:  
    *"Not in the past three months, one to three months ago"*
  - For `_p3m`:  
    *"Not in the past three months, Less than monthly"*


We used whether or not the activity occurred in the past month ("_pm") for cross-sectional analysis, but cohort analysis can use more granular data ("_p3m") and more variables.

```{r}
# activities_last_month <- data %>% select(starts_with("CONNECTION_activities") & ends_with("_pm")) %>% colnames()

activities_past_three_months<- data %>% dplyr::select(starts_with("CONNECTION_activities") & ends_with("_p3m")) %>% colnames()


activities_vars <- which_datasets_group(activities_past_three_months) %>% select(-cross22) %>% filter(if_all(2:7, ~ .x == "TRUE")) %>% select(variable_name) %>% unlist() %>% as.vector()
```

The issues with social number are similar.
'How many close friends do you have?' is represented by the variable 'CONNECTION_social_num_close_friends_grouped'
'Over the past three months, how many people belonging to each of the following categories have you spent time socializing with? Close friends' is represented by the variable 'CONNECTION_social_num_close_friends_p3m'

Other variables that are similar but do not ask the same are 'CONNECTION_social_num_close_relationships' and 
'CONNECTION_social_num_close_relationships_preference' asking about current and ideal number of friends and family members


```{r}
connections <- data %>% select(starts_with("CONNECTION_social_num")) %>% colnames() %>% sort()
# connections[connections %>% stringr::str_ends("_grouped")]

redundant_vars <- c("CONNECTION_social_num_close_friends_p3m", #exists as group
                    "CONNECTION_social_num_coworkers_and_classmates_p7d", #combines two categories but as raw data
                    "CONNECTION_social_num_coworkers_p7d",  #original variable
                    "CONNECTION_social_num_classmates_p7d", #original variable
                    "CONNECTION_social_num_family_p7d", #original variable
                    "CONNECTION_social_num_friends_p7d", #original variable
                    "CONNECTION_social_num_neighbours_p7d") #original variable

social_num_var_use <- which_datasets_group(connections)
# social_num_var_use %>% filter(!variable_name %in% redundant_vars) %>% View()

# Variables in all datasets
number_of_social_connections_vars <- social_num_var_use %>% filter(!variable_name %in% redundant_vars) %>%
  filter(if_all(2:8, ~ .x == "TRUE")) %>% select(variable_name) %>% unlist() %>% as.vector()

number_of_social_connections_vars
```

CONNECTION_social_time asks 'In the PAST WEEK, how many hours in total did you spend socializing with others from the following groups?'
CONNECTION_social_days asks 'In the PAST WEEK, how many days did you spend at least 5 minutes socializing with people from the following groups?'

They are virtually the same question, however, the answers are not necessarily consistent.



```{r}
social_time <- data %>% select(starts_with("CONNECTION_social_time"), starts_with("CONNECTION_social_days")) %>% colnames() %>% sort()
# social_time[social_time %>% stringr::str_ends("_grouped")]
red_vars <- c("CONNECTION_social_days_coworkers_and_classmates_p7d","CONNECTION_social_days_coworkers_p7d","CONNECTION_social_days_classmates_p7d","CONNECTION_social_days_family_p7d","CONNECTION_social_days_friends_p7d","CONNECTION_social_days_neighbours_p7d")
red_vars <- c(red_vars, sub("_days_","_time_",red_vars))
  
social_time_var_use <- which_datasets_group(social_time)

# social_time_var_use %>% filter(!variable_name %in% red_vars) %>% View()

social_time_vars <- social_time_var_use %>% filter(!variable_name %in% red_vars) %>%
  filter(if_all(2:8, ~ .x == "TRUE")) %>% select(variable_name) %>% unlist() %>% as.vector()

social_time_vars

df_temp <- data[social_time_vars]
df_temp$consistent <- T
for(i in 1:4){
hours <-  social_time_vars[i+4]
days <-  social_time_vars[i]
df_temp$consistent[df_temp[days]=="None (0 Days)" & !is.na(df_temp[days]) & df_temp[hours] != "No time"] <- F
}

# 8.3% of respondents said they did not spend at least five minutes with one group and then reported at least 1 hour of time spent together in the last week.

round(prop.table(table(df_temp$consistent)),2)
```

```{r}
# Participants identified as sexual or gender minorities at baseline, subsequent waves asked about trans identity and sexual orientation direclty
which_datasets_group(c("DEMO_gender","DEMO_orientation","DEMO_gender_trans","DEMO_identity_lgbtq"))

data$LGBTQ <- rep(FALSE, nrow(data))

data$LGBTQ[data$DEMO_orientation %in% c("Gay or Lesbian", "Bisexual or Queer", "Asexual", "Questioning") & !is.na(data$DEMO_orientation)] <- TRUE
data$LGBTQ[data$DEMO_gender_trans=="Yes" & !is.na(data$DEMO_gender_trans)] <- TRUE
data$LGBTQ[data$DEMO_gender=="Non-binary" & !is.na(data$DEMO_gender)] <- TRUE
data$LGBTQ[data$DEMO_identity_lgbtq=="Sexual or gender minorities (e.g., LGBTQ2+)" & !is.na(data$DEMO_identity_lgbtq)] <- TRUE
data$LGBTQ[is.na(data$DEMO_orientation) & is.na(data$DEMO_gender_trans) & is.na(data$DEMO_gender) & is.na(data$DEMO_identity_lgbtq)] <- NA


```
Study 1. Cross-sectional data

```{r}

# Selecting cross sectional data and relevant variables, including demographic and geographic data
cross <- data %>% filter(DATASET %in% c("2021 Cross-Sectional", "2023 Cross-Sectional", "2024 Cross-Sectional")) %>% select(UNIQUE_id, LGBTQ,starts_with("DEMO_"),WORK_hours_per_week,starts_with("GEO_"),all_of(c(activities_vars,number_of_social_connections_vars, social_time_vars)), LONELY_dejong_emotional_loneliness_sub_scale_score,LONELY_dejong_social_loneliness_sub_scale_score,DATASET) %>% droplevels.data.frame()

# Deleting empty columns
cross <- cross %>% select(-all_of(names(cross)[colSums(is.na(cross)) == nrow(cross)]))

# Identifying columns that do not appear in all waves 
miss_cols <- c()
for(i in levels(cross$DATASET)){
  df <- cross %>% filter(DATASET==i)
all_na_cols <- names(df)[colSums(is.na(df)) == nrow(df)]  
miss_cols <- c(miss_cols,all_na_cols)
}

# Deleting columns that do not appear in all cross sections
cross <- cross %>% select(-all_of(miss_cols)) %>% select(-starts_with("GEO_commute"), -ends_with("_y_n"))


# 2240 participants (19.4%) did not provide information on gender or sexual orientation
# table(cross$LGBTQ,useNA = "always")
# prop.table(table(cross$LGBTQ,useNA = "always"))
```

```{r}
levels(cross$DEMO_ethnicity)[levels(cross$DEMO_ethnicity) %in% c("Chinese", "Japanese", "Korean")] <- "East Asian"
levels(cross$DEMO_ethnicity)[levels(cross$DEMO_ethnicity) %in% c("Filipino", "Southeast Asian (e.g., Vietnamese, Cambodian, Laotian, Thai, etc.)")] <- "Southeast Asian"
levels(cross$DEMO_ethnicity)[levels(cross$DEMO_ethnicity) %in% c("South Asian (e.g., East Indian, Pakistani, Sri Lankan, etc.)", "West Asian (e.g., Iranian, Afghan, etc.)")] <- "South Asian and West Asian"
levels(cross$DEMO_ethnicity)[levels(cross$DEMO_ethnicity) %in% c("Arab", "None of the above")] <- "Other"

levels(cross$DEMO_household_income)[21:25] <- "$150,000 to $199,999"
levels(cross$DEMO_household_income)[16:20] <- "$100,000 to $149,999" 
levels(cross$DEMO_household_income)[12:15] <- "$60,000 to $99,999"
levels(cross$DEMO_household_income)[9:11] <- "$40,000 to $59,999"
levels(cross$DEMO_household_income)[5:8] <- "$20,000 to $39,999"
levels(cross$DEMO_household_income)[1:4] <- "Under $20,000"

cross <- cross %>% mutate(WORK_hours_per_week=factor(case_when(
  WORK_hours_per_week==0 ~ "Unemployed",
  WORK_hours_per_week<20 ~ "Works <20 h/week",
  WORK_hours_per_week<=168 ~ "Works 20 h/week or more",
  TRUE ~ NA
)))

# Creating categorical variables for age
cross <- cross %>%
  mutate(DEMO_age_group = factor(cut(
    DEMO_age,
    breaks = c(16, 20, 25, 30, 35, 40, 45, 50, 55, 60, 65, Inf),
    labels = c(
      "16–19", "20–24", "25–29", "30–34", "35–39",
      "40–44", "45–49", "50–54", "55–59", "60–64", "65+"
    ),
    right = FALSE
  ), ordered = T)) %>% relocate(DEMO_age_group, .after = DEMO_age)


cross <- cross %>% mutate(DEMO_age_group_2=factor(case_when(
  DEMO_age<30 ~ "< 30 years",
  DEMO_age %in% c(30:64) ~ "30 to 64 years",
  DEMO_age %in% c(65:100) ~ "65 and older",
  TRUE ~ NA
))) %>% relocate(DEMO_age_group_2, .after = DEMO_age_group)

levels(cross$DEMO_orientation)[levels(cross$DEMO_orientation) %in% c("Asexual", "Questioning", "None")] <- "None of the above"

levels(cross$DEMO_relationship_status)[levels(cross$DEMO_relationship_status)=="Single and dating"] <- "In a relationship"

levels(cross$GEO_province)[levels(cross$GEO_province) %in% c("Yukon", "Northwest Territories", "Nunavut")] <- "Territories"

cross <- cross %>% mutate(GEO_housing_household_size=factor(case_when(is.na(GEO_housing_household_size) ~ NA_character_,
                                                                GEO_housing_household_size==0 ~ "Alone",
                                                                GEO_housing_household_size==1 ~ "1 other person",
                                                                TRUE ~ "2 other people or more"),levels = c("Alone", "1 other person", "2 other people or more"), ordered = T))

cross <- cross %>% dplyr::mutate(DEMO_gender_so = factor(case_when(
  DEMO_gender=="Non-binary" ~ "Non-binary",
  LGBTQ==T & DEMO_gender=="Woman"~ "2SLGBTQ+ Women",
  LGBTQ==T & DEMO_gender=="Man" ~ "2SLGBTQ+ Men",
  LGBTQ==F & DEMO_gender=="Man" ~ "Cis-Het Men",
  LGBTQ==F & DEMO_gender=="Woman" ~ "Cis-Het Women",
  TRUE ~ NA))) %>% select(-DEMO_gender,-DEMO_orientation, -LGBTQ) %>% relocate(DEMO_gender_so, .before = DEMO_ethnicity) %>% droplevels.data.frame()

```


```{r}
demo_vars <- colnames(cross)[c(2,5:14)] #contains 3 variables for age
# Looking at differences between those with and without missing data
cross <- cross %>% mutate(missing_outcome =  is.na(LONELY_dejong_emotional_loneliness_sub_scale_score) | is.na(LONELY_dejong_social_loneliness_sub_scale_score)) 

missing_data_table <- cross %>% CreateTableOne(vars = demo_vars, strata = "missing_outcome", data = ., includeNA = F)

missing_data_table
tab_csv <- print(missing_data_table, printToggle = FALSE)
write.csv(tab_csv, file = here("cscs","patterns_social_connection","cscs_cross_2025_missing_outcome.csv"))



```

```{r}
# Creating  the analytic file
df <- cross
df[df=="Presented but no response"] <- NA
# df <-df %>% filter(missing_exposure==F & missing_outcome==F) %>% droplevels.data.frame() %>% select(-starts_with("missing_"),-GEO_residence_canada) 

df <-df %>% filter(missing_outcome==F) %>% droplevels.data.frame() %>% select(-starts_with("missing_"),-GEO_residence_canada) #GEO_residence_canada is excluded due to all being Canadian
```

Below we see the characteristics of participants missing outcome information, they are significantly younger that those who did provide outcome data, more likely to identify as "Man", ]to be single and dating, to identify as White, to be less educated, less likely to work full-time (provincial differences not reported).

```{r}
descr_tab <- df %>% CreateTableOne(vars = c(demo_vars, "LONELY_dejong_emotional_loneliness_sub_scale_score",
"LONELY_dejong_social_loneliness_sub_scale_score"), strata = "DATASET", data = ., includeNA = F, test = F, addOverall = T)

tab_1_csv <- print(descr_tab, printToggle = FALSE)
write.csv(tab_1_csv, file = here("cscs","patterns_social_connection","cscs_cross_2025_t1.csv"))
```


# Dealing with Missing Data 
```{r}

# Proportion missing
missing_info <- DataExplorer::profile_missing(df) 
#View(missing_info)

#Little's test
cross %>% select(starts_with("CONNECTION_"), LONELY_dejong_emotional_loneliness_sub_scale_score, LONELY_dejong_social_loneliness_sub_scale_score) %>% naniar::mcar_test() #NMAR
df %>% select(starts_with("CONNECTION_")) %>% naniar::mcar_test() #NMAR

# Transforming to integer due to very high OOBE with ordered factors

vars_to_num <- df %>% select(starts_with("CONNECTION_")) %>% colnames()
df[vars_to_num] <- lapply(df[vars_to_num],as.integer)
df[vars_to_num] <- df[vars_to_num] - 1 #Zero-ing "None"

# Listwise deletion
demo_vars <- demo_vars[demo_vars !="GEO_residence_canada"]
ids_complete <- df %>% select(-all_of(demo_vars)) %>%  filter(complete.cases(.)) %>% select(UNIQUE_id) %>% unlist() %>%  as.vector()
df_complete_cases <- df %>% filter(UNIQUE_id %in% ids_complete)

# Imputation
if(file.exists(file = here("cscs","patterns_social_connection","cscs_cross_2025_imputed.RDS"))){
  df_imputed <- readRDS(here("cscs","patterns_social_connection","cscs_cross_2025_imputed.RDS"))} else {
    
    p_load(parallel)
    p_load(doParallel)  
    p_load(foreach) 
    total_cores <- detectCores()
    used_cores <- floor(0.75 * total_cores)
    cl <- makeCluster(used_cores)
    registerDoParallel(cl)
    
    df_temp <- df %>% select(-DEMO_age_group,-DEMO_age_group_2, -UNIQUE_id)

    
    p_load(missForest)

    miss_df <- missForest(df_temp, verbose=T,parallelize = "forests", variablewise = T)
    
    imputation_info <- missing_info %>% filter(feature %in% colnames(df_temp))
    imputation_info$OOBE <- miss_df$OOBerror
    
    
    mse_1 <- mean(imputation_info$OOBE[which(colnames(df_temp) %in% activities_vars)]) #1 to 7
    mse_2 <- mean(imputation_info$OOBE[which(colnames(df_temp) %in% c(number_of_social_connections_vars,social_time_vars[1:4]))]) #1 to 4
    mse_3 <- mean(imputation_info$OOBE[which(colnames(df_temp) %in% social_time_vars[5:8])]) #1 to 3

    nrmse_1 <- sqrt(mse_1)/6
    nrmse_2 <- sqrt(mse_2)/3
    nrmse_3 <- sqrt(mse_3)/2    
    
    
    weighted_nrmse <- (18 * nrmse_1 + 9 * nrmse_2 + 4 *nrmse_3) / 31
    weighted_nrmse

    df_imputed <- miss_df$ximp
    df_imputed[c("DEMO_age_group", "DEMO_age_group_2")] <- df[c("DEMO_age_group", "DEMO_age_group_2")]
    saveRDS(df_imputed, file = here("cscs","patterns_social_connection","cscs_cross_2025_imputed.RDS"))
    stopCluster(cl)
  }
patterns <- c(number_of_social_connections_vars, social_time_vars)
df_imputed[c(activities_vars,patterns)] <- lapply(df_imputed[c(activities_vars,patterns)], round)
```


# Reducing data dimensionality

```{r}
p_load("psych")

parallel_analysis_results <- fa.parallel(df_imputed[activities_vars], fa = "fa", n.iter = 100)

factor_analysis_results <- fa(df_imputed[activities_vars], nfactors = parallel_analysis_results$nfact, rotate = "oblimin", fm = "ml")

print(factor_analysis_results$loadings, cutoff = 0.2) 

factor_analysis_scores <- factor_analysis_results$scores %>% as.data.frame() %>% 
  rename(group_activities = ML1, 
         closed_meetings = ML4, 
         physical_intimacy = ML2, 
         digital_activities = ML5,
         phone_conversation = ML3,
         ) 

df_imputed[colnames(factor_analysis_scores)] <- factor_analysis_scores

parallel_analysis_patterns <- fa.parallel(df_imputed[patterns], fa = "fa", n.iter = 100)
fa_pattern_results <- fa(df_imputed[patterns], nfactors = parallel_analysis_patterns$nfact, rotate = "oblimin", fm = "ml")

print(fa_pattern_results$loadings, cutoff = 0.2) 

pattern_scores <- fa_pattern_results$scores %>% as.data.frame() %>% 
  rename(time_colleagues = ML1, 
         time_neighbours = ML2, 
         time_family = ML3,
         time_friends = ML4)

df_imputed[colnames(pattern_scores)] <- pattern_scores

```
Creating a simplified numeric dataset
```{r}
df_simplified <- df_imputed %>% select(all_of(c(colnames(pattern_scores), colnames(factor_analysis_scores))), ends_with("sub_scale_score"))

##################  COMPLETE CASES  ##################
# df_simplified_cc <- df_complete_cases %>% select(all_of(colnames(df_simplified)))
######################################################

colnames(df_simplified) <- sub("LONELY_dejong_","",colnames(df_simplified))
colnames(df_simplified) <- sub("_sub_scale_score","",colnames(df_simplified))

##################  COMPLETE CASES  ##################
# colnames(df_simplified_cc) <- colnames(df_simplified)
# corp_cc <- plot_correlation(df_simplified_cc)
######################################################

corp <- DataExplorer::plot_correlation(df_simplified)
p_load(ggplot2)
ggsave(here("cscs","patterns_social_connection","cscs_cross_2025_correlation_plot.png"), plot = corp, width = 2000, height = 2000, units = "px", dpi = 300)

p_load(bnlearn)

network_structure_result <- boot.strength(df_simplified, R=500, algorithm = "hc") 
avg_bn <- averaged.network(network_structure_result, threshold = 0.85)
avg_bn_95 <- averaged.network(network_structure_result, threshold = 0.95)

##################  COMPLETE CASES  ##################
# cc_boot <- boot.strength(df_simplified_cc, R=500, algorithm = "hc") 
# cc_avg_bn <- averaged.network(cc_boot, threshold = 0.85)
######################################################

##################  COMBINED DATASETS  ##################
# cc_boot_all <- rbind(cc_boot, strenght)
# avg_all <- averaged.network(cc_boot_all, threshold = 0.85)
# graphviz.plot(avg_bn)
# graphviz.plot(cc_avg_bn)
# graphviz.plot(avg_all)
########################################################

fitted_model <- bn.fit(avg_bn,data = df_simplified)
arcs_neg <- matrix(ncol=2)
colnames(arcs_neg) <- c("Var1", "Var2")
for(i in c(1:ncol(df_simplified))){
  node <- fitted_model[[i]]$node
  coeff <- fitted_model[[i]]$coefficients[fitted_model[[i]]$coefficients <0]
  
x <- expand.grid(names(coeff)[names(coeff) != "(Intercept)"],node)
arcs_neg <- rbind(arcs_neg,x)
}
arcs_neg <- arcs_neg[-1,]
colnames(arcs_neg) <- c("from", "to")
p_load(BiocManager)
p_load(Rgraphviz)


png(here("cscs","patterns_social_connection","cscs_cross_2025_dag_plot.png"), width = 800, height = 800, units = "px", res = 300)
graphviz.plot(fitted_model, highlight = list(arcs=as.matrix(arcs_neg), nodes=c("emotional_loneliness", "social_loneliness")))
dev.off()

fitted_model$emotional_loneliness
fitted_model$social_loneliness

```


```{r}
outcomes <- c("LONELY_dejong_emotional_loneliness_sub_scale_score","LONELY_dejong_social_loneliness_sub_scale_score")

p_load(factoextra)
fviz_nbclust(cbind(factor_analysis_scores,pattern_scores), kmeans, method = "wss")  #no clear elbow
fviz_nbclust(cbind(factor_analysis_scores,pattern_scores), kmeans, method = "silhouette") #2 activity clusters

cluster_social <- kmeans(cbind(factor_analysis_scores,pattern_scores), centers = 2, nstart = 25)


df_imputed$cluster_social <- as.factor(as.character(cluster_social$cluster))
CreateTableOne(vars=c(colnames(factor_analysis_scores),colnames(pattern_scores), outcomes), strata = "cluster_social", data=df_imputed)

```


```{r}
df_factor <- df_imputed %>% select(DEMO_age_group_2, DEMO_gender_so,DEMO_household_income, DEMO_ethnicity, GEO_housing_household_size, starts_with("cluster"),all_of(outcomes))

colnames(df_factor) <- sub("DEMO_","",colnames(df_factor))
colnames(df_factor) <- sub("LONELY_dejong_","",colnames(df_factor))
colnames(df_factor) <- sub("_sub_scale_score","",colnames(df_factor))

df_factor <- df_factor %>% rename(social_prescription=cluster_social)
levels(df_factor$social_prescription) <- c("Yes", "No")
 
black_list_vars <- expand.grid(from=colnames(df_factor),to=c("age_group_2", "gender_so","ethnicity")) %>% as.data.frame()

df_factor[1:ncol(df_factor)] <- lapply(df_factor[1:ncol(df_factor)], as.character)
df_factor[1:ncol(df_factor)] <- lapply(df_factor[1:ncol(df_factor)], as.factor)
demo_dag <- boot.strength(df_factor, R=500, algorithm = "hc", algorithm.args = list(blacklist = black_list_vars))
avg_demo <- averaged.network(demo_dag, threshold = 0.85)

fitted_model_demo <- bn.fit(avg_demo, data = df_factor)
png(here("cscs","patterns_social_connection","cscs_cross_2025_dag_factor_plot.png"), width = 800, height = 800, units = "px", res = 300)
graphviz.plot(fitted_model_demo)
dev.off()

emo_lone <- as.data.frame(as.table(fitted_model_demo$emotional_loneliness$prob))

fig_dep <- ggplot(emo_lone, aes(x = age_group_2, y = Freq, fill = emotional_loneliness)) +
    geom_bar(stat = "identity", position = "fill") +
    scale_y_continuous(labels = scales::percent_format()) +
    labs(
        title = "Conditional Probabilities of Emotional Loneliness",
        x = "Age group",
        y = "Proportion",
        fill = "Emotional Loneliness"
    ) +
    theme_minimal() +
    scale_fill_brewer(palette = "Set2")
ggsave(here("cscs","patterns_social_connection","cscs_cross_2025_probs_emo_lon.png"), plot = fig_dep, width = 2000, height = 2000, units = "px", dpi = 300)

df_dag <- as.data.frame(as.table(fitted_model_demo$social_loneliness$prob))
fig_dependencies <- ggplot(df_dag, aes(x = emotional_loneliness, y = Freq, fill = social_loneliness)) +
    geom_bar(stat = "identity", position = "fill") +
    facet_wrap(~ social_prescription, labeller = label_both) +
    scale_y_continuous(labels = scales::percent_format()) +
    labs(
        title = "Conditional probabilities of Social Loneliness levels by levels of emotional loneliness and social prescribing",
        x = "Emotional Loneliness",
        y = "Proportion",
        fill = "Social Loneliness"
    ) +
    theme_minimal() +
    scale_fill_brewer(palette = "Set2")
ggsave(here("cscs","patterns_social_connection","cscs_cross_2025_probs_soc_lon.png"), plot = fig_dependencies, width = 2000, height = 2000, units = "px", dpi = 300)
```


```{r}
bn.fit.barchart(fitted_model_demo$social_prescription)

bn.fit.dotplot(fitted_model_demo$social_prescription, main = "Conditional probabilities of social prescription", xlab = "Age group", ylab = "Gender and sexual orientation")

bn.fit.dotplot(fitted_model_demo$social_loneliness, xlab = "Social prescription", ylab = "Emotional loneliness", main = "Conditional probabilities of social loneliness")

```
```{r}
list_act <- df_imputed %>% select(starts_with("CONNECTION_act")) %>% purrr::map(table) 

for(i in 1:18)
  name_activity <-sub("_p3m","", sub("CONNECTION_activities_","",names(list_act[1])))
  activities_count <- data.frame(activity=name_activity,
  group=levels(cross[[names(list_act[1])]])[levels(cross[[names(list_act[1])]]) != "Presented but no response"],
  total=as.vector(list_act[[1]]))
  
for(i in 2:18){
  name_activity <-sub("_p3m","", sub("CONNECTION_activities_","",names(list_act[i])))
  activities_count <- rbind(activities_count, data.frame(activity=name_activity,
  group=levels(cross[[names(list_act[i])]])[levels(cross[[names(list_act[i])]]) != "Presented but no response"],
  total=as.vector(list_act[[i]])))
}
p_load(tidyverse)
p_load(showtext)
p_load(ggtext)
p_load(waffle)
p_load(MetBrewer)

font_add_google("Outfit", "title_font")
font_add_google("Cabin", "body_font")
showtext_auto()

title_font <- "title_font"
body_font <- "body_font"

title_text <- "Frequency of social activities in the past three months"
subtitle_text <- "Source: Canadian Social Connection Survey"
caption_text <- "Healthy Ecologies and Lifestyles Lab, Simon Fraser University"

activities_count$group <- factor(activities_count$group, levels = unique(activities_count$group))
activities_count$activity <- factor(activities_count$activity)
levels(activities_count$activity) <- c("temple","coffee","PC game","discuss","gym","zoom", "help","hug", "kiss","meeting","befriend","phone","sex","SMS","visit_fam","visit_pal","volunteer","walk")

activities_count$total_adjusted <- round(activities_count$total/10,digits = 0)
df_graph <- activities_count %>% filter(group!="Not in the past three months") %>% droplevels.data.frame()

p = ggplot(df_graph, aes(fill = group, values = total_adjusted)) +
  geom_waffle(color = "white", size = .25, n_rows = 10, flip = TRUE) +
  facet_wrap(~activity, nrow = 1, strip.position = "bottom") +
  scale_x_discrete() +
  scale_y_continuous(breaks = c(0, 20, 40, 60),
                     labels = c("0", "2000", "4000", "6000"))+
  MetBrewer::scale_fill_met_d("Hiroshige", direction=1)+
  coord_equal()+
  labs(title = title_text,
       subtitle = subtitle_text,
       caption = caption_text)+
  theme_minimal()+
  theme(
    axis.title = element_blank(),
    axis.text.x = element_text(family = body_font, size = 12, angle = 90, hjust = 1, vjust = 0.5),
    axis.text.y = element_text(family = body_font, size = 12),
  # Legend
  legend.position = "top",
  legend.title = element_blank(),
  legend.spacing = unit(0.5, 'cm'),
  legend.key.height= unit(0.5, 'cm'),
  legend.key.width= unit(0.7, 'cm'),
  legend.text = element_text(family = body_font,
                             size=13,
                             face = 'plain',
                             color = "grey10"),
  
  # TITLE
  plot.title.position = "plot",
  plot.title = element_textbox(margin = margin(20, 0, 10, 0),
                               size = 30,
                               family = title_font,
                               face = "bold",
                               width = unit(55, "lines")),
  
  # SUB-TITLE
  plot.subtitle = element_text(margin = margin(10, 0, 20, 0),
                            size = 16,
                            family = body_font,
                            color = "grey15"),
  # Caption
  plot.caption = element_text(family=body_font,
                              face="plain",
                              size=14, 
                              color="grey40",
                              hjust=.5,
                              margin=margin(20,0,0,0)),
  
  plot.background = element_rect(color="white", fill="white"),
  plot.margin = margin(20, 40, 20, 40)
) +theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))

showtext_opts(dpi = 320)
ggsave(
  here("cscs","patterns_social_connection","cscs_cross_2025_waffle.png"),
  dpi=320, width = 12, height = 9
)
showtext_auto(FALSE)
```

```{r}

rdf2 <- data %>%
  filter(UNIQUE_id %in% cross$UNIQUE_id) %>%
  select(all_of(number_of_social_connections_vars)) %>%
  mutate(across(everything(), as.character))  %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "level") %>%
  count(variable, level, name = "n") 

rdf2 <- rdf2 %>% filter(!is.na(level) & level != "Presented but no response") 
rdf2$level[grepl("none",rdf2$level, ignore.case = T)] <- "0 (None)"
rdf2$level[grepl("1",rdf2$level, ignore.case = T)] <- "1 or 2"
rdf2$level[grepl("3",rdf2$level, ignore.case = T)] <- "3 or 4"  
rdf2$level[grepl("5",rdf2$level, ignore.case = T)] <- "5 or more"  
rdf2$variable[grepl("close_frie",rdf2$variable, ignore.case = T)] <- "Close friends" 
rdf2$variable[grepl("family",rdf2$variable, ignore.case = T)] <- "Family" 
rdf2$variable[grepl("neighb",rdf2$variable, ignore.case = T)] <- "Neighbours" 
rdf2$variable[grepl("coworkers",rdf2$variable, ignore.case = T)] <- "Colleagues" 
rdf2$variable[grepl("grouped",rdf2$variable, ignore.case = T)] <- "Friends" 


rdf2<-  rdf2 %>%
  group_by(variable) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()

rdf2$n[rdf2$level=="0 (None)"] <- -1*rdf2$n[rdf2$level=="0 (None)"]
rdf2$prop[rdf2$level=="0 (None)"] <- -1*rdf2$prop[rdf2$level=="0 (None)"]

p_load(hrbrthemes)

rdf2 <- rdf2 %>%
  mutate(variable = factor(variable, levels = c("Close friends", "Friends", "Family", "Neighbours", "Colleagues"),ordered = T)) %>%
  arrange(variable)

rdf2$level <- factor(rdf2$level, levels = c("0 (None)", "5 or more", "3 or 4", "1 or 2"), ordered = T)

ggplot(rdf2, aes(fill=level, y=prop, x=variable)) + 
  geom_bar(position="stack", stat="identity") + 
  scale_fill_met_d("Hiroshige", d=1) +  theme_ipsum() +
  xlab("Social group") + 
  ylab("Proportion") +
  ggtitle("People you spent at least 5 minutes with during the past week") +
    scale_y_continuous(labels = scales::percent_format()) + theme_bw()
```
```{r}
efa_copy <- factor_analysis_results
rownames(efa_copy$loadings) <-  c("temple","coffee","PC game","discuss","gym","zoom", "help","hug", "kiss","meeting","befriend","phone","sex","SMS","visit_fam","visit_pal","volunteer","walk")
psych::fa.diagram(efa_copy)
```
```{r}
p_load(ggstatsplot)
p_load(patchwork)

# Generate list of plots
plots_activities <- lapply(colnames(factor_analysis_scores), function(y) {
  ggbetweenstats(
    data = df_imputed,
    x = cluster_social,
    y = !!sym(y),
    xlab = "Social prescription",
    ylab = stringr::str_to_sentence(sub("_"," ",y)),
    title = ""
    ) 
})

# Combine them in a 2x2 grid
combined_plot <- (plots_activities[[1]] | plots_activities[[2]]) /
                 (plots_activities[[3]] | plots_activities[[4]])

#combined_plot

plots_patterns <- lapply(colnames(pattern_scores), function(y) {
  ggbetweenstats(
    data = df_imputed,
    x = cluster_social,
    y = !!sym(y),
    xlab = "Social prescription",
    ylab = stringr::str_to_sentence(sub("_"," ",y)),
    title = ""
    ) 
})

combined_plot_pat <- (plots_patterns[[1]] | plots_patterns[[2]]) /
                 (plots_patterns[[3]] | plots_patterns[[4]])

#combined_plot_pat
```

